#Analysis of Harmful and Damaging Evens in Storm Data

## Preparation of The Data
At the very beginning, we need to read the data from the data file and set the plotting environment by include the library of "ggplot". 
```{r}
library(ggplot2)
setwd("~/Desktop/my_documents/coursera/fall_2015/reproducible_research/project_2")
data=read.csv('./repdata-data-StormData.csv',sep=',',header=TRUE)
```

Since there is no explicity information on the damages of properties and crops, we need to add two new features into the data about these information. 
```{r}
data$prop.damage <- with(data, PROPDMG * ifelse(PROPDMGEXP == "", 1, ifelse(PROPDMGEXP == 
"K", 1000, ifelse(PROPDMGEXP == "M", 1e+06, ifelse(PROPDMGEXP == "B", 1e+09,-1)))))

data$crop.damage <- with(data, CROPDMG * ifelse(CROPDMGEXP == "", 1, ifelse(CROPDMGEXP == 
"K", 1000, ifelse(CROPDMGEXP == "M", 1e+06, ifelse(CROPDMGEXP == "B", 1e+09,-1)))))
```

## Analysis of the Harm to Population Health
In this section we analyze the most harmful events to population health. According to the information provided by the data, we divide the analysis into two parts: fatal analysis and injury analysis. 
### Fatal Analysis
```{r}
s.fatal=split(data$FATALITIES,as.factor(data$EVTYPE))
fatal.sum=sapply(s.fatal,sum)
fatal.sum.sorted=fatal.sum[order(fatal.sum,decreasing=TRUE)]
fatal.mean=sapply(s.fatal,mean)
fatal.mean.sorted=fatal.mean[order(fatal.sum,decreasing=TRUE)]
event.fatal.sorted=names(fatal.sum.sorted)
summary.fatal=data.frame(event.type=event.fatal.sorted,total.fatal.number=fatal.sum.sorted,
                         fatal.per.event=fatal.mean.sorted,row.names=NULL)
```
By the code above, we can see the ten most harmful events according to fatal in the table below: 
```{r}
print(summary.fatal[1:10,])
```
We can also plot the distribution: 
```{r fig.width=16}
data.fatal.top10=subset(data,data$EVTYPE %in% summary.fatal[1:10,1])

ggplot(data.fatal.top10,aes(x=as.factor(EVTYPE),y=FATALITIES))+geom_boxplot()+
  ggtitle("Distribution of Fatalities for the most Harmful Events")+ylab("log10 of the number of fatals")+
  xlab("Event Type") + scale_y_log10() + theme_bw()
```

### Injury Analysis
```{r}
s.injur=split(data$INJURIES,as.factor(data$EVTYPE))
injur.sum=sapply(s.injur,sum)
injur.sum.sorted=injur.sum[order(injur.sum,decreasing=TRUE)]
injur.mean=sapply(s.injur,mean)
injur.mean.sorted=injur.mean[order(injur.sum,decreasing=TRUE)]
event.injur.sorted=names(injur.sum.sorted)
summary.injur=data.frame(event.type=event.injur.sorted,total.injury.number=injur.sum.sorted,
                         injuries.per.event=injur.mean.sorted,row.names=NULL)
```
By the code above, we can see the ten most harmful events according to fatal in the table below: 
```{r}
print(summary.injur[1:10,])
```
We can also plot the distribution: 
```{r fig.width=16}
data.injur.top10=subset(data,data$EVTYPE %in% summary.injur[1:10,1])

ggplot(data.injur.top10,aes(x=as.factor(EVTYPE),y=INJURIES))+geom_boxplot()+
  ggtitle("Distribution of Injuries for the most Harmful Events")+ylab("log10 of the number of injuries")+
  xlab("Event Type") + scale_y_log10() + theme_bw()
```


## Analysis of Harm to Economic Damage
In this section we analyze the most harmful events to economic damage. According to the information provided by the data, we divide the analysis into two parts: property damage analysis and crop damage analysis. 
### Property Damage Analysis
```{r}
s.prop=split(data$prop.damage,as.factor(data$EVTYPE))
prop.sum=sapply(s.prop,sum)
prop.sum.sorted=prop.sum[order(prop.sum,decreasing=TRUE)]
prop.mean=sapply(s.prop,mean)
prop.mean.sorted=prop.mean[order(prop.sum,decreasing=TRUE)]
event.prop.sorted=names(prop.sum.sorted)
summary.prop=data.frame(event.type=event.prop.sorted,total.prop.damage.number=prop.sum.sorted,
                        prop.damage.per.event=prop.mean.sorted,row.names=NULL)
```
By the code above, we can see the ten most harmful events according to property damage in the table below: 
```{r}
print(summary.prop[1:10,])
```
We can also plot the distribution: 
```{r fig.width=16}
data.prop.top10=subset(data,data$EVTYPE %in% summary.prop[1:10,1])

ggplot(data.prop.top10,aes(x=as.factor(EVTYPE),y=prop.damage))+geom_boxplot()+
  ggtitle("Distribution of Properties Damage for the most Harmful Events")+ylab("log10 of the number of properties damage")+
  xlab("Event Type") + theme_bw()+scale_y_log10()
```

### Crop Damage Analysis
```{r}
s.crop=split(data$crop.damage,as.factor(data$EVTYPE))
crop.sum=sapply(s.crop,sum)
crop.sum.sorted=crop.sum[order(crop.sum,decreasing=TRUE)]
crop.mean=sapply(s.crop,mean)
crop.mean.sorted=crop.mean[order(crop.sum,decreasing=TRUE)]
event.crop.sorted=names(crop.sum.sorted)
summary.crop=data.frame(event.type=event.crop.sorted,total.crop.damage.number=crop.sum.sorted,
                        crop.damage.per.event=crop.mean.sorted,row.names=NULL)
```
By the code above, we can see the ten most harmful events according to property damage in the table below: 
```{r}
print(summary.crop[1:10,])
```
We can also plot the distribution: 
```{r fig.width=16}
data.crop.top10=subset(data,data$EVTYPE %in% summary.crop[1:10,1])

ggplot(data.crop.top10,aes(x=as.factor(EVTYPE),y=crop.damage))+geom_boxplot()+
  ggtitle("Distribution of crops Damage for the most Harmful Events")+ylab("log10 of the number of crops damage")+
  xlab("Event Type") + theme_bw()+scale_y_log10()
```


